Review of Edge Detectors

### 2014-11-11 old_versions/edge-detection-review-v7.tex
second round of third revision, improving references

%-------------------------------------------------------------------------------
% A Review of Classic Edge Detectors - v8
% by Haldo Spontón & Juan Cardelino
% IPOL 2012
%-------------------------------------------------------------------------------

@@

\begin{abstract}
In this paper some of the classic alternatives for edge detection in digital images are studied. The main idea 
behind edge detection is to find where abrupt changes in the intensity of an image have occurred. 
The first family of algorithms reviewed in this work uses the first derivative to find the changes of intensity, 
such as Sobel, Prewitt and Roberts. In the second family reviewed, second derivative is used, for example in algorithms 
like Marr-Hildreth and Haralick. \\
Results obtained from a qualitative point of view (perceptual) and from a quantitative 
point of view (number of operations, execution time) are compared, considering different ways to convolve an 
image with a kernel (step required in some of the algorithms).
\end{abstract}

@@

Results of the implemented algorithms are presented in section \ref{sec:results}, along with examples 
to compare their performance. Conclusions are detailed in section \ref{sec:conclusions}.

@@

To compute the gradient, the computation of partial derivatives $\partial f/\partial x$ and $\partial f/\partial y$ 
at every pixel in the image is required. When dealing with digital images, numerical approximations 
of the partial derivatives over a neighborhood about a point are required. In the following,
the methods of Roberts, Prewitt and Sobel will be studied, whose main difference is how they perform this calculation.

@@

The most usual approach to approximate the first derivative is to take the Taylor expansion of first order with
a small $h$. Thus $f'(x)$ is computed as
\begin{equation}
	f'(x) \simeq \frac{f(x+h) - f(x)}{h}.\\
\end{equation}
The image is composed by discrete points of coordinates $(i,j)$, so taking $x=i*h$ and $y=j*h$, the 
first derivative of the image is approximated, based on the intensity values ​​at points of the image, as:
\begin{equation}
\label{eq:roberts1}
	g_x = \frac{\partial f(x,y)}{\partial x} = f(i+1,j) - f(i,j)
\end{equation}
and
\begin{equation}
\label{eq:roberts2}
	g_y = \frac{\partial f(x,y)}{\partial y} = f(i,j+1) - f(i,j).\\
\end{equation}

Equations \ref{eq:roberts1} and \ref{eq:roberts2} can be implemented for all values of $x$ and $y$
by filtering the image $f(x,y)$ with the 1-D masks in Figure \ref{fig:1dmasks}. But when finding 
diagonal edges is desired, 2-D masks are needed. The \textit{Roberts operators} are one of the earliest 
attempts to use 2-D masks for this purpose. These operators are based on computing the diagonal 
diferences implemented by filtering an image with the masks in Figure \ref{fig:roberts}. By convention 
in these figures, x-coordinates grow from left to right and y-coordinates grow top-down. It is also 
usual to normalize the masks, in order to have gain equal to $1$. \\

@@

Masks of size $2\times2$, although conceptually simple, are not symmetrical with respect to the central points. 
Having symmetrical edges is a desirable property and can only be achieved with oddly sized masks, the smallest 
of them being the 3x3. These masks provide more information to find the direction of the edges, because they take 
into account information on opposite sides of the central point. \\

@@

These masks can be obtained (beyond normalization) as the convolution of a horizontal derivation mask 
$[-1\ 0\ 1]$ with a vertically moving average $[1\ 1\ 1]$, and vice versa. Hence, these operators have 
smoothing properties in the direction opposite to the gradient.

@@

A slight variation of the Prewitt operators uses more weight on the central coefficients of the 
difference. It can be shown that using a $2$ in the center location provides better image smoothing. This
variation is implemented using masks in Figure \ref{fig:sobel}. These operators are called 
\textit{Sobel operators}. \\

@@

Sobel masks can be seen (beyond normalization) as the convolution of a horizontal derivation mask 
$[-1\ 0\ 1]$ with a vertical smoothing filter $[1\ 2\ 1]$ (closer to a Gaussian response than Prewitt), and vice versa. 
Hence, these operators have better smoothing properties, as mentioned in the preceding paragraph. \\

The computational difference between Prewitt and Sobel masks is not an issue. It is preferable to 
use Sobel operators, because they are better localized, and their filter are also less aliased, 
because of the weighted shape of $[1\ 2\ 1]$.

@@

\begin{figure}[b!]
	\centering
	\subfigure[$th = 0.1*\max{M}$.]{\includegraphics[width=0.3\textwidth]{thresholded_image_1.png}}
	\quad
	\subfigure[$th = 0.2*\max{M}$.]{\includegraphics[width=0.3\textwidth]{thresholded_image_2.png}}
	\quad
	\subfigure[$th = 0.4*\max{M}$.]{\includegraphics[width=0.3\textwidth]{thresholded_image_3.png}}
	\caption{Thresholded images. Different values of the threshold applied to the magnitude of the gradient.}
	\label{fig:thresholding}
\end{figure}

Then, once the respective operators are applied, an approximation of the gradient (stored in two images) 
is obtained, containing the approximation of the partial derivatives $g_x$ and $g_y$. Gradient magnitude 
image $M$ is calculated using the equation \ref{eq:mag_alpha} (Figure \ref{fig:profiles3d-c}). \\

Finally, the edges image is obteined by thresholding the gradient magnitude image, i.e.\ for each pixel in the image compute
\begin{equation*}
	\mbox{edges}[i] = \begin{cases} 0,& \mbox{if}\ M[i]\geq\mbox{threshold} \\
									1,& \mbox{if}\ M[i]<\mbox{threshold}.
					  \end{cases}
\end{equation*}
A black and white image is obtained, in which edge points are indicated in white (see Figure \ref{fig:thresholding}). 
It is noted that different thresholds lead to different results. Thicker edges for smaller thresholds, lack of edges 
for larger thresholds. This threshold, and the others mentioned below, are defined as a percentage of the maximum value 
of the image to be processed. The advantage of this is to adapt the threshold to the dynamic range of the image. 
Therefore, the parameter $threshold$ in the algorithms take values ​​between $0$ and $1$.

@@

The pseudo-code of implemented algorithms is shown in Algorithm \ref{algo:fded}.
\clearpage

@@

The Marr-Hildreth algorithm is a method of detecting edges in digital 
images. It is based on finding zero crossing points of the second derivative
of the image. This can be done in several ways. Two different ways of doing 
this are implemented implemented in this work (see block diagram in Figure 
\ref{fig:blockdiagram2}); convolving the image with a Gaussian kernel and then 
approximating the second derivative (Laplacian) with a 3x3 kernel, or 
convolving the image with a kernel calculated as the Laplacian of a 
Gaussian function. There are more ways to do so, for example, using 
recursive Gaussian filters \cite{Deriche1993Recursively}. \\
\clearpage

@@

Zero crossing detection is the key feature of the Marr-Hildreth edge detection method. The technique 
presented in the previous paragraph is attractive for its simplicity of implementation and its low 
computational cost. In general it is a technique that yields good results, but if more precision in 
finding the zero crossings is needed, more advanced methods for finding zero crossings with subpixel 
accuracy could be used (e.g.\ marching squares \cite{marching_cubes}).

@@

The pseudo-code of Marr-Hildreth's algorithm is shown in Algorithm \ref{algo:mh}.

@@

In this section the original work of Haralick \cite{bb20239} on edge detection is presented in detail.
The main idea behind this algorithm is identical to that of the previous method; find zeros in 
the second derivative of the image. In this method, the input image is smoothly approximated through local bi-cubic
polynomial fitting. Then, when calculating the second derivative analytically, it is possible to find 
an equivalent expression to find the zeros of the second derivative of the polynomial as a function of 
its parameters.%\footnote{This implementation is slightly different from the traditional implementation 
%of the Haralick algorithm. We do not use the condition concerning the third derivative (to be negative), 
%in order to implement two edge detection methods using only the second derivative (along the 
%Marr-Hildreth algorithm).}

@@

Here, the interpolation method used by Haralick is presented, although there are other options to do this \cite{getreuer}. \\

@@

Consider 25 points in a small neighborhood of a point $(x,y)$ in the image. This gives us an equal number of equations 
to find 10 coefficients. As mentioned before this leads to an overdetermied system which can be solved by least squares. 
By substituting the 25 data points into the polinomial equation (\ref{eq:bicubic}), the following system of 
equations is obtained,

@@

Using matrix notation, the system can be rewritten as

@@

The neighborhood of each point of the image is approximated using the bi-cubic polynomial 
expression in equation \ref{eq:bicubic}. If just the first order terms of this 
polynomial are taken, the gradient angle $\theta$, defined with negative y-axis, can be approxmated as

\begin{align}
\label{eq:sincos}
	\sin(\theta) & = -\frac{k_2}{\sqrt{k_2^2 + k_3^2}} \nonumber \\
	\cos(\theta) & = -\frac{k_3}{\sqrt{k_2^2 + k_3^2}}. \nonumber \\
\end{align}

@@

\begin{equation*}
	x = \rho\cos{\theta} \: \text{and} \: y = \rho\sin{\theta} \\
\end{equation*}

@@

\begin{align}
	f'_{\theta}(\rho) = C_1 + 2C_2\rho + 3C_3\rho^2 \nonumber \\
	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho \nonumber \\
	f'''_{\theta}(\rho) = 6C_3 .\nonumber \\
\end{align}

@@

\begin{equation}
	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho = 0 \ \ \Rightarrow \ \ \left| \frac{C_2}{3C_3} \right| < \rho_0,
\end{equation}

where $\rho_0$ is a fixed threshold, passed as an argument to the algorithm. \\

The other condition required is that the third derivative to be negative, i.e.\

\begin{equation}
	f'''_{\theta}(\rho) = 6C_3 < 0 \ \ \Rightarrow \ \ C_3 < 0.
\end{equation}

@@

\begin{enumerate}
	\item Find the coefficients $k_1 \hdots k_{10}$, as shown in section \ref{sec:bicubic}.
	\item Compute $\sin(\theta)$ and $\cos(\theta)$ (equations \ref{eq:sincos}).
	\item Compute $C_2$ and $C_3$ (equations \ref{eq:c}).
	\item If $\left| \frac{C_2}{3C_3} \right| < \rho_0$ and $C_3 < 0$, then that point is an edge point.
\end{enumerate}

@@

The pseudo-code of Haralick's algorithm is shown in Algorithm \ref{algo:haralick}.

@@

All algorithms implemented use some common mathematical operations that are independent of the algorithms themselves. 
These operations, although basic, have a great impact on the algorithm outcome, and thus they need to be 
implemented with care. In this section those operations are reviewed.

@@

The first step in every algorithm presented above is to convert color images to gray intensity images. 
For this purpose, \href{http://www.libpng.org/}{libpng} coefficients are used,
\begin{equation}
    Y = (6968 R + 23434 G + 2366 B) / 32768
\end{equation}
where $R$, $G$ and $B$ are the red, green and blue components respectively\footnote{Must be carefully implemented, using \texttt{float}. See code in section \ref{app:marr-hildreth}.}.

@@

The size of the kernel $n$ and the standard deviation of the exponential function $\sigma$ are both 
input parameters, but these are not strictly independent of each other. 
A value of $n$ big enough is needed to ensure that no information is lost when creating the kernel $G$. To ensure this, $n$ is taken equal to the first odd integer greater than $6\sigma$. Larger values of $n$ do not add more significant samples ​​of $G$, and increase the number of operations in the convolution.\\

@@

The Laplacian of Gaussian $\nabla^2G(x,y)$ can be obtained analyically first and then a discrete mask 
can be computed by sampling the analytical kernel. Using this kernel for edge detection involves only 
one convolution with the input image (unlike the case of Gaussian kernel, in which two convolutions 
have to be performed, one with the kernel and another with the Laplacian operator).\\

@@

When making a convolution, it is necessary to define the boundary conditions used to calculate such convolution 
around the edges of the image, and to get a valid output the same size as input image. In this paper, 
two methods were implemented: zero-padding and boundary reflection. Zero-padding implies completing 
the borders of the image with many zero valued pixels as needed (depending on the size of the convolution 
kernel). Reflection involves completing those pixels with the corresponding symmetric pixel value 
relative to the edge of the image. Convolution is not the only way of filtering an image with a kernel. There are other methods such as FFT convolution 
or recursive filtering. Implementations of these methods can be found in the \href{https://tools.ipol.im/wiki/author/code/hatchery/}{IPOL Code Hatchery}.

@@

This code documentation was generated using a perl script to extract some parts of the 
code and comments into latex files. This code documentation tool is available as a GIT
project: 
\begin{center} 
\url{https://github.com/juan-cardelino/source_comment_extractor}. \\
\end{center}
Also a Doxygen generated documentation of the functions used is available from:
\begin{center}
\url{http://iie.fing.edu.uy/~haldos/ipol/red_v0.1} \\
\end{center}
Required external libraries: 
\begin{center}
\texttt{iio} (\url{http://dev.ipol.im/git/coco/iio.git}) \\
\texttt{libpng} (\url{http://www.libpng.org/})
\end{center}

@@

The results of each algorithm are first shown in a simple case, where a test image composed of a white square on a black background (shown in Figure \ref{fig:original1}) is used. \\

@@

Figures \ref{fig:result1-a} to \ref{fig:result1-f} show the output of each algorithm. Execution times\footnote{Running on Intel Core i3 CPU (2.53GHz), 3 Gb RAM, standard laptop.} for each algorithm are shown in Table \ref{exectime1}. \\

@@

	\subfigure[Marr-Hildreth (Gaussian). $\sigma=1.5$, $n=13$, $th_{ZC}=0.1$.]{\label{fig:result1-d}\includegraphics[width=0.2\textwidth]{results/square127_marr-hildreth.png}}
	\quad
	\subfigure[Marr-Hildreth (LoG). $\sigma=1.5$, $n=17$, $th_{ZC}=0.1$.]{\label{fig:result1-e}\includegraphics[width=0.2\textwidth]{results/square127_marr-hildreth-log.png}}
	\quad
	\subfigure[Haralick. $\rho=0.4$.]{\label{fig:result1-f}\includegraphics[width=0.2\textwidth]{results/square127_haralick.png}}
	\caption{Results of the algorithms using the Figure \ref{fig:original1} as input image. The first three (\ref{fig:result1-a}, \ref{fig:result1-b} and \ref{fig:result1-c}) correspond to the first derivative methods (Roberts, Prewitt and Sobel), then the following two (\ref{fig:result1-d} and \ref{fig:result1-e}) are from the Marr-Hildreth algorithm using Gaussian and LoG kernels, and the last one (\ref{fig:result1-f}) corresponds to the Haralick algorithm.}
	\label{fig:result1}
\end{figure}

\begin{table}[t!]
	\begin{center}
	\begin{tabular}{| l | r |}
		\hline \rule{0pt}{3ex}
		\cellcolor[gray]{0.8} \textbf{Algorithm}	& \cellcolor[gray]{0.8} \textbf{Execution time (s)}	\\ \hline \rule{0pt}{3ex}
		Roberts, Prewitt and Sobel					& $0.020 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (Gaussian)					& $0.030 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (LoG)							& $0.050 \ s$										\\ \hline \rule{0pt}{3ex}
		Haralick									& $0.040 \ s$										\\
		\hline
	\end{tabular}
	\end{center}
	\caption{Execution time of the algorithms (including i/o) using the Figure \ref{fig:original1} as input image (128$\times$128 pixels, 1 channel).}
	\label{exectime1}
\end{table}
\vspace{0.5cm}

It is appreciated that in the simplest cases, the first derivative edge detection algorithms are running better and faster. All algorithms show consistent results. Thicker edges appear in the results of Haralick, due to smooth edges model that is imposed in this algorithm. \\

@@

\begin{figure}[t!]
	\centering
	\subfigure[Original image.]{\includegraphics[width=0.48\textwidth]{results/molino.jpg}}
	\quad
	\subfigure[Image converted to greyscale.]{\includegraphics[width=0.48\textwidth]{results/molino_bw.png}}
	\caption{Test image 2: windmill (1000$\times$563 pixels).}
	\label{fig:original2}
\end{figure}

@@

\begin{table}[h!]
	\begin{center}
	\begin{tabular}{| l | r |}
		\hline \rule{0pt}{3ex}
		\cellcolor[gray]{0.8} \textbf{Algorithm}	& \cellcolor[gray]{0.8} \textbf{Execution time (s)}	\\ \hline \rule{0pt}{3ex}
		Roberts, Prewitt and Sobel					& $0.550 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (Gaussian)					& $1.050 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (LoG)							& $1.440 \ s$										\\ \hline \rule{0pt}{3ex}
		Haralick									& $0.930 \ s$										\\
		\hline
	\end{tabular}
	\end{center}
	\caption{Execution time of the algorithms (including i/o) using Figure \ref{fig:original2} as input image (1000$\times$563 pixels, 3 channels).}
	\label{exectime2}
\end{table}

In this example the difference in performance between the first derivative edge detection algorithms and the second derivative ones is more significative. It can be seen in Figure \ref{fig:result3} an area of ​​interest of the image, enlarged to have a better view of the detail. Note that none of the first derivative methods (even with a loose threshold) detect the lower edge of the blade (which is shaded). Neither the Haralick algorithm is able to detect it. However, the Marr-Hildreth algorithm detects it, using either a Gaussian or a LoG kernel. \\

@@

Some observations: 
\begin{itemize}
	\item The first derivative algorithms, although they work properly and quickly, have some problems like discontinuity of the edges. They are also very affected by noise in images, appearing unwanted isolated edges (More sophisticated methods help to avoid and improve this, e.g. Canny edge detector \cite{Canny1986}). \\
	\item The Marr-Hildreth algorithm (in its two versions) achieved interesting results in the edge detail. This algorithm provides a first step to regularity in the image, since the filtering with a Gaussian kernel. This behavior can be seen in Lena's hair (Figure \ref{fig:example1}), or inside the oranges (example \ref{fig:example2}). \\
	\item As mentioned earlier, Haralick's algorithm is the first to assume greater regularity of the image in the neighborhood of a pixel (third order). This causes thicker edges, as shown in both examples. But, note that some edges are only detected with this algorithm, e.g. the lower edges of the oranges, which are in the shadow, in Figure \ref{fig:example2}. These edges are quite smooth, so that the Haralick model is well suited to them, and they do not represent an abrupt change in the intensity to be detected by other methods. \\
\end{itemize}

@@

Some of the most traditional methods of edge detection in digital images were discussed and carefully implemented in this work. The implemented algorithms were tested with synthetic and real images, obtaining generally the expected results, taking into account the limitations of these methods. \\

>>>>

%-------------------------------------------------------------------------------
% A Review of Classic Edge Detectors - v7
% by Haldo Spontón & Juan Cardelino
% IPOL 2012
%-------------------------------------------------------------------------------

@@

\begin{abstract}
In this paper some of the classic alternatives for edge detection in digital images are studied. The main idea 
behind edge detection is to find where abrupt changes in the intensity of an image have occurred. 
The first family of algorithms reviews the use of first derivative to find the changes of intensity, 
such as Sobel, Prewitt and Roberts. In the second family reviewed, second derivative is used, for example in algorithms 
like Marr-Hildreth and Haralick. \\
Results obtained from a qualitative point of view (perceptual) and from a quantitative 
point of view (number of operations, execution time) are compared, considering different forms to convolve an 
image with a kernel (step required in some of the algorithms).
\end{abstract}

@@

Results of the implemented algorithms are presented in section \ref{sec:results}, along some examples 
to compare their performance. Conclusions are detailed in section \ref{sec:conclusions}.

@@

The computation of partial derivatives $\partial f/\partial x$ and $\partial f/\partial y$ 
at every pixel in the image is required. When dealing with digital images, numerical approximations 
of the partial derivatives over a neighborhood about a point are required. In the following,
the methods of Roberts, Prewitt and Sobel will be studied, whose main difference is how they perform this calculation.

@@

The most usual approach of the first derivative is to take the Taylor expansion of first order with
a small $h$. Approximating the first derivative
\begin{equation}
	f'(x) \simeq \frac{f(x+h) - f(x)}{h}.\\
\end{equation}
The image has discrete points of coordinates $(i,j)$, so taking $x=i*h$ and $y=j*h$, the 
first derivative of the image is approximated, based on the intensity values ​​at points of the image, as:
\begin{equation}
\label{eq:roberts1}
	g_x = \frac{\partial f(x,y)}{\partial x} = f(i+1,j) - f(i,j)
\end{equation}
and
\begin{equation}
\label{eq:roberts2}
	g_y = \frac{\partial f(x,y)}{\partial y} = f(i,j+1) - f(i,j).\\
\end{equation}

Equations \ref{eq:roberts1} and \ref{eq:roberts2} can be implemented for all values of $x$ and $y$
by filtering the image $f(x,y)$ with the 1-D masks in Figure \ref{fig:1dmasks}. But when finding 
diagonal edges is desired, 2-D masks are needed. The \textit{Roberts operators} are one of the earliest 
attempts to use 2-D masks for this purpose. These operators are based on computing the diagonal 
diferences implemented by filtering an image with the masks in Figure \ref{fig:roberts}. By convention 
in these figures, x-coordinates grow from left to right and y-coordinates grow top-down. \\

@@

Masks of size $2\times2$, although conceptually simple, are not symmetrical with respect to the central points. 
Having symmetrical edges is a desirable property and can only be achieved with oddly sized mask, the smallest 
of them being the 3x3. These masks provide more information to find the direction of the edges, because they take 
into account information on opposite sides of the central point. \\

@@

These masks can be obtained (beyond normalization) as the convolution of a horizontal derivation mask 
$[-1\ 0\ 1]$ with vertically moving average $[1\ 1\ 1]$, and vice versa. Hence, these operators have 
smoothing properties.

@@

A slight variation of the Prewitt operators uses more weight on the central coefficients of the 
difference. It can be shown that using a $2$ in the center location provides image smoothing. This
variation is implemented using masks in Figure \ref{fig:sobel}. These operators are called 
\textit{Sobel operators}. \\

@@

Sobel masks can be obtained (beyond normalization) as the convolution of a horizontal derivation mask 
$[-1\ 0\ 1]$ with vertically filter $[1\ 2\ 1]$ (closer to a Gaussian response than Prewitt), and vice versa. 
Hence, these operators have smoothing properties, as mentioned in the preceding paragraph. \\

Computational difference between Prewitt and Sobel masks is not an issue. It is preferable to 
use Sobel operators, because they are better localized, and their filter are also less aliased, 
because of the Gaussian shape of $[1\ 2\ 1]$.

@@

\begin{figure}[b!]
	\centering
	\subfigure[$th = 0.1*\max{M}$.]{\includegraphics[width=0.3\textwidth]{thresholded_image_1.png}}
	\quad
	\subfigure[$th = 0.2*\max{M}$.]{\includegraphics[width=0.3\textwidth]{thresholded_image_2.png}}
	\quad
	\subfigure[$th = 0.4*\max{M}$.]{\includegraphics[width=0.3\textwidth]{thresholded_image_3.png}}
	\caption{Thresholded images.}
	\label{fig:thresholding}
\end{figure}

Then, once the respective operators are applied, an approximation of the gradient (stored in two images) 
is obtained, containing the approximation of horizontal and vertical components of the first derivative, 
$g_x$ and $g_y$ respectively. Gradient magnitude image $M$ is calculated using the equation \ref{eq:mag_alpha} 
(Figure \ref{fig:profiles3d-c}). \\

Finally, edges image is obteined by thresholding the gradient magnitude image, i.e.\ for each pixel in the image compute
\begin{equation*}
	\mbox{edges}[i] = \begin{cases} 0,& \mbox{if}\ M[i]\geq\mbox{threshold} \\
									1,& \mbox{if}\ M[i]<\mbox{threshold}.
					  \end{cases}
\end{equation*}
A black and white image is obtained, in which edge points are indicated in white (see Figure \ref{fig:thresholding}). 
It is noted that different thresholds lead to different results. Thicker edges for smaller thresholds, lack of edges 
for larger thresholds.

@@

Pseudo-code of implemented algorithms is shown in Algorithm \ref{algo:fded}.
\clearpage

@@

The Marr-Hildreth algorithm is a method of detecting edges in digital 
images. It is based on find zero crossing points of the second derivative
of the image. This can be done in several ways. Two different ways of doing 
this are implemented implemented in this work (see block diagram in Figure 
\ref{fig:blockdiagram2}); convolving the image with a Gaussian kernel and then 
approximating the second derivative (Laplacian) with a 3x3 kernel, or 
convolving the image with a kernel calculated as the Laplacian of a 
Gaussian function. There are more ways to do so, for example, using 
recursive Gaussian filters \cite{Deriche1993Recursively}. \\
\clearpage

@@

Zero crossing detection is the key feature of the Marr-Hildreth edge detection method. The technique 
presented in the previous paragraph is attractive for its simplicity of implementation and its low 
computational cost. In general it is a technique that yields good results, but if more precision in 
finding the zero crossings is needed, methods for finding zero crossings with subpixel accuracy 
should be used (e.g.\ marching squares \cite{marching_cubes}).

@@

Pseudo-code of Marr-Hildreth's algorithm is shown in Algorithm \ref{algo:mh}.

@@

This section explains in detail the original work by Haralick, particularly is applied the same 
technique bicubic polynomial fitting (see \ref{sec:bicubic}). There are other options for image 
interpolation \cite{getreuer}. \\

The idea behind the Haralick edge detector is identical to that of the previous method; find zeros in 
the second derivative of the image. In this method, the input image is smoothed through local bi-cubic
polynomial fitting. Then, when calculating the second derivative analytically, it is possible to find 
an equivalent expression to find the zeros of the second derivative of the polynomial as a function of 
its parameters.\footnote{This implementation is slightly different from the traditional implementation 
of the Haralick algorithm. We do not use the condition concerning the third derivative (to be negative), 
in order to implement two edge detection methods using only the second derivative (along the 
Marr-Hildreth algorithm).}

@@

@@

Consider 25 points in a small neighborhood of a point $(x,y)$ in the image. This gives us an equal number of equations 
to find 10 coefficients. As mentioned before this leads to an overdetermied system which can be solved by least squares. 
By substituting the 25 data points into the polinomial equation (equation \ref{eq:bicubic}), the following system of 
equations is obtained,

@@

Using matrix notation,

@@

The neighborhood of each point of the image is approximated using the bi-cubic polynomial 
expression in equation \ref{eq:bicubic}. If just the first order terms of this 
polynomial are taken, the gradient angle $\theta$, defined with positive y-axis, is defined as

\begin{align}
\label{eq:sincos}
	\sin(\theta) & = \frac{k_2}{\sqrt{k_2^2 + k_3^2}} \nonumber \\
	\cos(\theta) & = \frac{k_3}{\sqrt{k_2^2 + k_3^2}}. \nonumber \\
\end{align}

@@

\begin{equation*}
	x = \rho\cos{\theta} \: \text{and} \: y = \rho\sin(\theta) \\
\end{equation*}

@@

\begin{align}
	f'_{\theta}(\rho) = C_1 + 2C_2\rho + 3C_3\rho^2 \nonumber \\
	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho .\nonumber \\
\end{align}

@@

\begin{equation}
	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho = 0 \ \ \Rightarrow \ \ \left| \frac{C_2}{3C_3} \right| < \rho_0 .
\end{equation}

@@

\begin{enumerate}
	\item Find the coefficients $k_1 \hdots k_{10}$, as shown in section \ref{sec:bicubic}.
	\item Compute $\sin(\theta)$ and $\cos(\theta)$ (equations \ref{eq:sincos}).
	\item Compute $C_2$ and $C_3$ (equations \ref{eq:c}).
	\item If $\left| \frac{C_2}{3C_3} \right| < \rho_0$, then that point is an edge point.
\end{enumerate}

@@

Pseudo-code of Haralick's algorithm is shown in Algorithm \ref{algo:haralick}.

@@

All algorithms implemented use some common mathematical operations that are independent of the algorithms themselves. 
These operations, although basic, have a great impact on the algorithm outcome, and thus they need to be 
implemented with care.

@@

The first step in every algorithm presented above is to convert color images to gray intensity images. 
For this purpose, \href{http://www.libpng.org/}{libpng} coefficients are used,
\begin{equation}
    Y = (6968 R + 23434 G + 2366 B) / 32768
\end{equation}
where $R$, $G$ and $B$ are the red, green and blue components respectively.

@@

The size of the kernel $n$ and the standard deviation of the exponential function $\sigma$ are both 
input parameters, but these are not strictly independent of each other. 
A large enough value of $n$ is needed to ensure that no information is lost when creating the kernel $G$. To ensure this, $n$ is taken equal to the first odd integer greater than $6\sigma$. Larger values of $n$ do not add more significant samples ​​of $G$, and increase the number of operations in the convolution.\\

@@

The Laplacian of Gaussian $\nabla^2G(x,y)$ can be obtained analyically first and then a discrete mask 
can be computed by sampling the analytical kernel. Using this kernel for edge detection involves only 
one convolution with the input image (unlike the case of Gaussian kernel, in which two convolutions 
have to be maded, one with the kernel and another with the Laplacian operator).\\

@@

When making a convolution, it is necessary to define the boundary conditions used to calculate such convolution 
around the edges of the image, and to get a valid output the same size as input image. In this paper, 
two methods were implemented: zero-padding and boundary reflection. Zero-padding implies completing 
the borders of the image with many zero valued pixels as needed (depending on the size of the convolution 
kernel). Reflection involves completing those pixels with the corresponding symmetric pixel value 
relative to the edge of the image. \\

@@

This code documentation was generated using a perl script to extract some parts of the 
code and comments into latex files. This code documentation tool is available as a GIT
project: 
\begin{center} 
\url{https://github.com/juan-cardelino/source_comment_extractor}. \\
\end{center}
Also a Doxygen generated documentation of the functions used is available from:
\begin{center}
\url{http://iie.fing.edu.uy/~haldos/ipol/red_v0.1} \\
\end{center}
Required external libraries: \texttt{iio} (\url{http://dev.ipol.im/git/coco/iio.git}).

@@

First the results of each algorithm are shown, using a simple test input image, the white square on black background shown in Figure \ref{fig:original1}. \\

@@

Figures \ref{fig:result1-a} to \ref{fig:result1-f} show the output of each algorithm. The first three correspond to the first derivative methods (Roberts, Prewitt and Sobel), then the following two are from the Marr-Hildreth algorithm using Gaussian and LoG kernels, and the last one corresponds to the Haralick algorithm. Execution times\footnote{Running on Intel Core i3 CPU (2.53GHz), 3 Gb RAM, standard laptop.} for each algorithm are shown in Table \ref{exectime1}. \\

@@

	\subfigure[Marr-Hildreth (Gaussian). $\sigma=1.5$, $n=13$, $th_{ZC}=0.1$.]{\label{fig:result1-d}\includegraphics[width=0.2\textwidth]{results/square127_marr-hildreth.png}}
	\quad
	\subfigure[Marr-Hildreth (LoG). $\sigma=1.5$, $n=17$, $th_{ZC}=0.1$.]{\label{fig:result1-e}\includegraphics[width=0.2\textwidth]{results/square127_marr-hildreth-log.png}}
	\quad
	\subfigure[Haralick. $\rho=0.4$.]{\label{fig:result1-f}\includegraphics[width=0.2\textwidth]{results/square127_haralick.png}}
	\caption{Results of the algorithms using the Figure \ref{fig:original1} as input image.}
	\label{fig:result1}
\end{figure}

\begin{table}[t!]
	\begin{center}
	\begin{tabular}{| l | r |}
		\hline \rule{0pt}{3ex}
		\cellcolor[gray]{0.8} \textbf{Algorithm}	& \cellcolor[gray]{0.8} \textbf{Execution time (s)}	\\ \hline \rule{0pt}{3ex}
		Roberts, Prewitt and Sobel					& $0.020 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (Gaussian)					& $0.030 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (LoG)							& $0.050 \ s$										\\ \hline \rule{0pt}{3ex}
		Haralick									& $0.040 \ s$										\\
		\hline
	\end{tabular}
	\end{center}
	\caption{Execution time of the algorithms using the Figure \ref{fig:original1} as input image (128$\times$128 pixels, 1 channel).}
	\label{exectime1}
\end{table}
\vspace{0.5cm}

It is appreciated that in the simplest cases, the first derivative edge detection algorithms are running better and faster. All algorithms show consistent results, but the Haralick algorithm shows some rare artifacts out of the edges. This may occur because this is the only algorithm that assumes a certain regularity (third degree) in the neighborhood of a pixel. Synthetic edges are far from this regularity. This may cause the adjustment of the bicubic polynomial function around these points to be quite different from reality. \\

@@

\begin{figure}[t!]
	\centering
	\subfigure[Original image.]{\includegraphics[width=0.48\textwidth]{results/molino.jpg}}
	\quad
	\subfigure[Grayscale image.]{\includegraphics[width=0.48\textwidth]{results/molino_bw.png}}
	\caption{Test image 2: windmill (1000$\times$563 pixels).}
	\label{fig:original2}
\end{figure}

@@

\begin{table}[h!]
	\begin{center}
	\begin{tabular}{| l | r |}
		\hline \rule{0pt}{3ex}
		\cellcolor[gray]{0.8} \textbf{Algorithm}	& \cellcolor[gray]{0.8} \textbf{Execution time (s)}	\\ \hline \rule{0pt}{3ex}
		Roberts, Prewitt and Sobel					& $0.550 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (Gaussian)					& $1.050 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (LoG)							& $1.440 \ s$										\\ \hline \rule{0pt}{3ex}
		Haralick									& $0.930 \ s$										\\
		\hline
	\end{tabular}
	\end{center}
	\caption{Execution time of the algorithms using Figure \ref{fig:original2} as input image (1000$\times$563 pixels, 3 channels).}
	\label{exectime2}
\end{table}

In this example the difference in performance between the first derivative edge detection algorithms and the second derivative ones is more significative. It can be seen in Figure \ref{fig:result3} an area of ​​interest of the image, enlarged to have a better view of the detail. Note that none of the first derivative methods (even with a loose threshold) detect the lower edge of the blade (which is shaded). Neither the Haralick algorithm detects it. However, the Marr-Hildreth algorithm detects it, using either a Gaussian or a LoG kernel. \\

@@

Some observations: 
\begin{itemize}
	\item The first derivative algorithms, although they work properly and quickly, have some problems like discontinuity in the edges. They are also very affected by noise in images, appearing unwanted isolated edges (More sophisticated methods help to avoid and improve this, e.g. Canny edge detector \cite{Canny1986}). \\
	\item The Marr-Hildreth algorithm (in its two versions) achieved interesting results in the edge detail. This algorithm provides a first step to regularity in the image, since the filtering with a Gaussian kernel. This behavior can be seen in Lena's hair (Figure \ref{fig:example1}), or inside the oranges (example \ref{fig:example2}). \\
	\item As mentioned earlier, Haralick's algorithm is the first to assume greater regularity of the image in the neighborhood of a pixel (third order). This causes thicker edges, as shown in both examples. But, note that some edges are only detected with this algorithm, e.g. the lower edges of the oranges, which are in the shadow, in Figure \ref{fig:example2}. These edges are quite smooth, so that the Haralick model is well suited to them, and they do not represent an abrupt change in the intensity to be detected by other methods. \\
\end{itemize}

@@

Some of the most traditional methods of edge detection in digital images are discussed and carefully implemented in this work. The implemented algorithms were tested with synthetic and real images, obtaining generally the expected results, taking into account the limitations of these methods. \\

The first derivative algorithms (Roberts, Prewitt and Sobel) have the advantage of having a very simple implementation. Also they run extremely fast, because they only consists of a convolution with a very small kernel (2$\times$2 or 3$\times$3 pixels). The results obtained with these methods are quite good, considering their simplicity, but they have problems such as noise and discontinuity at the edges. \\
